\Title

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction

%\ninept
%
%\abstract{Abstract}
%
\keywords{\Keywords}
%
\maketitle
%%%%%%%%% BODY TEXT
\section{Introduction}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
\label{sec:intro}
% >>>>>>>>>>> LuP is here <<<<<<<<<<<
%
\subsection{Recent challenges in machine learning and computer vision}
%%%-----------------------------------------------------------------
Machine learning, and computer vision in particular, have undergone a major transformation in modern times, leaving science fiction movies to become mainstream tools found in the smartphones in everyone's pocket. Less visible, but even more crucial to our economy and security, are the many automated tools that now dominate the range of services that equip virtually all sectors of our daily lives. This trend is most pronounced in the field of medicine, as it provides practitioners with new diagnostic tools to treat people faster and more effectively. This particular example also illustrates some important challenges that these tools are increasingly facing. One of these is their sensitivity to adversarial attacks, that is, the sensitivity of the models' output to minute changes in the input. This could raise important questions about the reliability of the results of in any automated diagnostic model. More generally, these models do not explain why they produce a given output for a class of inputs. This lack of explainability is a hindrance to the advancement of these methodologies as well as to their acceptance for most safety-relevant tasks. If we take a step back, we will remember that these tools are inherited from theoretical models designed to understand the brain. While the first artificial neural networks were mainly applied to " toy " type problems, they also offered a growing number of applications in computer vision, but also more generally for task automation. In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates, and today they reach higher performances than humans for image categorization. One of the challenges of these models is to be able to scale them up for future applications, for example in autonomous driving. Indeed, in such applications, the input does not consist of a single image (say a few million pixels) but of a continuous stream of images from several cameras.

@@

Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights. Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, hetero-synaptic delays. In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay in the computational process. As a consequence, this extends our previous model by encoding spiking motifs as 3D images. The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier. In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these hetero-synaptic delays. %

@@

In this work, we study the emergence of such spatiotemporal spike patterns when training a single layer of spike neurons on a supervised classification task (\seeFig{motion_task}). The paper is organized as follows. We develop an SNN-like method capable of learning hetero-synaptic delays to perform motion detection on an event-based synthetic dataset.   We will first detail the methodology by defining the core spike mechanism using hetero-synaptic delays. We will then formally define the computational task of motion detection. This will allow us to formalize the final spiking neuron used to learn the model and test its effectiveness.  In the results section, we will detail the emergent structure of the model solving the task and show in particular the similarities with neurobiological observations. We will also test the network with classical stimuli used in neuroscience, and in particular the frequency or contrast responses.  We will also study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning. This will allow us to show how such a model can provide an effective solution to the energy/accuracy trade-off.  Finally, we will conclude by highlighting the main contributions of this paper, while defining some limitations. This will open up perspectives for future SNNs.  As neuromorphic devices are, by design, good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is transferable to this type of hardware.

>>>>

%\ninept
%
%\abstract{Abstract}
%
\keywords{\Keywords}
%
\maketitle
%%%%%%%%% BODY TEXT
\section{Introduction}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
\label{sec:intro}
% >>>>>>>>>>> LuP is here <<<<<<<<<<<
%
\subsection{Recent challenges in machine learning and computer vision}
%%%-----------------------------------------------------------------
Machine learning, and computer vision in particular, have undergone a major transformation in modern times, leaving science fiction movies to become mainstream tools found in the smartphones in everyone's pocket. Less visible, but even more crucial to our economy and security, are the many automated tools that now dominate the range of services that equip virtually all sectors of our daily lives. This trend is most pronounced in the field of medicine, as it provides practitioners with new diagnostic tools to treat people faster and more effectively. This particular example also illustrates some important challenges that these tools are increasingly facing. One of these is their sensitivity to adversarial attacks, that is, the sensitivity of the models' output to minute changes in the input. This could raise important questions about the reliability of the results of in any automated diagnostic model. More generally, these models do not explain why they produce a given output for a class of inputs. This lack of explainability is a hindrance to the advancement of these methodologies as well as to their acceptance for most safety-relevant tasks. If we take a step back, we will remember that these tools are inherited from theoretical models designed to understand the brain. While the first artificial neural networks were mainly applied to toy-like problems, they also offered a growing number of applications in computer vision, but also more generally for task automation. In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates, and today they reach higher performances than humans for image categorization. One of the challenges of these models is to be able to scale them up for future applications, for example in autonomous driving. Indeed, in such applications, the input does not consist of a single image (say a few million pixels) but of a continuous stream of images from several cameras.

@@

This paper explores the possibility that spiking motifs may be a useful representation used in neurobiology and that we may transsfer this to neuromorphic algorithms. Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights. Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, hetero-synaptic delays. In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay in the computational process. As a consequence, this extends our previous model by encoding spiking motifs as 3D images. The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier. In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these hetero-synaptic delays. %

@@

In this work, we study the emergence of such spatiotemporal spike motifs when training a single layer of spike neurons on a supervised classification task (\seeFig{motion_task}). The paper is organized as follows. We develop an SNN-like method capable of learning hetero-synaptic delays to perform motion detection on an event-based synthetic dataset.   We will first detail the methodology by defining the core spike mechanism using hetero-synaptic delays. We will then formally define the computational task of motion detection. This will allow us to formalize the final spiking neuron used to learn the model and test its effectiveness.  In the results section, we will detail the emergent structure of the model solving the task and show in particular the similarities with neurobiological observations. We will also test the network with classical stimuli used in neuroscience, and in particular the frequency or contrast responses.  We will also study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning. This will allow us to show how such a model can provide an effective solution to the energy/accuracy trade-off.  Finally, we will conclude by highlighting the main contributions of this paper, while defining some limitations. This will open up perspectives for future SNNs.  As neuromorphic devices are, by design, good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is transferable to this type of hardware.
